---
title: "PEV and genetic variance effects in the accuracia calculation"
author: "Marcio Araujo"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    code_folding: hide
    toc: no
    toc_depth: '2'
    df_print: paged
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: no
    fig_caption: yes
    toc-location: left
  html_notebook:
    code_folding: hide
    toc: yes
    toc_depth: '2'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: no
    fig_caption: yes
    toc-location: left
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F
)
```

## Background

In quantitative genetics, selective accuracy is an important metric that measures the reliability of the estimated breeding values (EBVs) or genotypic values (GVs). It quantifies how closely the estimated values reflect the true genetic merit of individuals. This is particularly useful in breeding programs where individuals are selected based on their genetic potential.

Selective accuracy (\(r\)) can be computed using the Prediction Error Variance (PEV) and the genetic variance (\(\sigma^2_g\)). The **PEV** represents the variance of the difference between the true and estimated genetic values, while the **genetic variance** refers to the variance of the breeding values in the population. The relationship between these variables is described by the following formula:

\[
r = \sqrt{1 - \frac{\text{PEV}}{\sigma^2_g}}
\]

Where:
- \(r\) is the selective accuracy.
- \(\text{PEV}\) is the Prediction Error Variance.
- \(\sigma^2_g\) is the genetic variance.

The selective accuracy increases as the PEV decreases or as the genetic variance increases, indicating more reliable estimates of genetic merit. 

In this document, we will explore how variations in PEV and genetic variance affect selective accuracy through simulation, using various scenarios for both parameters. This will help us understand the influence of these factors on the accuracy of selection decisions in breeding programs.



```{r pct}
# Load necessary packages
library(ggplot2)

```

```{r calculation}
# Function to calculate selective accuracy
calculate_accuracy <- function(pev, genetic_variance) {
  round(sqrt(1 - (pev / genetic_variance)), 2)
}

# Define ranges for genetic variance and PEV
genetic_variance_seq <- seq(50, 200, by = 25)  # Genetic variance ranging from 50 to 200
pev_seq <- seq(5, 100, by = 5)                # PEV ranging from 5 to 100

# Create an empty data frame to store the results
results <- expand.grid(pev = pev_seq, genetic_variance = genetic_variance_seq)
results$accuracy <- NA

# Calculate accuracy for each combination of PEV and genetic variance
for (i in 1:nrow(results)) {
  results$accuracy[i] <- calculate_accuracy(results$pev[i], results$genetic_variance[i])
}

```


```{r plot}
# Create a plot showing the effect of PEV and genetic variance on accuracy
plot <- ggplot(results, aes(x = pev, y = accuracy, color = as.factor(genetic_variance))) +
  geom_point() +
  geom_line(size = 1) +
  labs(
    title = "Effect of PEV and Genetic Variance on Selective Accuracy",
    x = "PEV (Prediction Error Variance)",
    y = "Selective Accuracy",
    color = "Genetic Variance"
  ) +
  geom_hline(
    yintercept = 0.8,
    linetype = "dashed",
    linewidth = 0.1
    
  ) +
  scale_color_viridis_d() +
  theme_classic() +
  guides(
    color = guide_legend(reverse = TRUE)
  )
  

plotly::ggplotly(plot)
```

```{r Final}

```

